\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-11 doc/USER_MANUAL/07_adjoint_simulations.tex

Adjoint simulations are generally performed for two distinct applications.
First, they can be used in point source moment-tensor inversions, or source imaging for
earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
Second, they can be used to generate finite-frequency sensitivity
kernels that are a critical part of tomographic inversions based upon
3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
case, source parameter or velocity structure updates are sought to
minimize a specific misfit function (e.g., waveform or traveltime
differences), and the adjoint simulation provides a means of computing
the gradient of the misfit function and further reducing it in successive
iterations. Applications and procedures pertaining to source studies
and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
 In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

@@

Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
solver.}{\small \par}
\begin{itemize}
\item Notice that attenuation is not implemented yet for the computation
of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
in the \texttt{Par\_file}.
\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
to avoid too much high-frequency noise in the forward wavefield, although
theoretically the high-frequency noise should be eliminated when convolved
with an adjoint wavefield with the proper frequency content.
\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
in the previous chapters in that the state variables for the last
time step of the simulation, including wavefields of the displacement,
velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
to be used for the subsequent simulation. }{\small \par}
\item For regional simulations, the files recording the absorbing boundary
contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
= .true.}.
\end{itemize}
\item \textbf{Prepare the adjoint sources}

@@

With the successful forward simulation and the adjoint source ready
in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
solver.
\begin{itemize}
\item The adjoint simulation is launched together with the back reconstruction
of the original forward wavefield from the state variables saved from
the previous forward simulation, and the finite-frequency kernels
are computed by the interaction of the reconstructed forward wavefield
and the adjoint wavefield.
\item The back-reconstructed seismograms at the original station locations
are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
and can be collected to the local disk.
\item These back-constructed seismograms can be compared with the time-reversed
original seismograms to assess the accuracy of the backward reconstruction,
and they should match very well.
\item The arrays for density, P-wave speed and S-wave speed kernels are
also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
are the different types of kernels.
\end{itemize}
\item \textbf{Run the anisotropic kernel simulation}

Instead of the kernels for the isotropic wave speeds, you can also
compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
(using Voigt's notation) of the elastic tensor in the cartesian coordinate
system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
\texttt{.true.} in \texttt{constants.h} before compiling the package.
The definition of the parameters $C_{IJ}$ in terms of the corresponding
components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
the processor number). The output kernels correspond to absolute perturbations
$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
For consistency, the output density kernels with this option turned
on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
anisotropic kernels can then be combined to obtain the kernels for
different parameterizations of anisotropy. This can be done, for example,
when combining the kernel files from slices into one mesh file (see
Section~\ref{sec:Finite-Frequency-Kernels}).

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
the package will save anisotropic kernels parameterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}

>>>>

Adjoint simulations are generally performed for two distinct applications.
First, they can be used in point source moment-tensor inversions, or source imaging for
earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
Second, they can be used to generate finite-frequency sensitivity
kernels that are a critical part of tomographic inversions based upon
3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
case, source parameter or velocity structure updates are sought to
minimize a specific misfit function (e.g., waveform or traveltime
differences), and the adjoint simulation provides a means of computing
the gradient of the misfit function and further reducing it in successive
iterations. Applications and procedures pertaining to source studies
and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
parameters in the \texttt{DATA/Par\_file} are \texttt{SIMULATION\_TYPE}
(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

@@

Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{DATA/Par\_file}{\small{}
(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
solver.}{\small \par}
\begin{itemize}
\item Notice that attenuation is not implemented yet for the computation
of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
in the \texttt{DATA/Par\_file}.
\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
to avoid too much high-frequency noise in the forward wavefield, although
theoretically the high-frequency noise should be eliminated when convolved
with an adjoint wavefield with the proper frequency content.
\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
in the previous chapters in that the state variables for the last
time step of the simulation, including wavefields of the displacement,
velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
to be used for the subsequent simulation. }{\small \par}
\item For regional simulations, the files recording the absorbing boundary
contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
= .true.}.
\end{itemize}
\item \textbf{Prepare the adjoint sources}

@@

With the successful forward simulation and the adjoint source ready
in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{DATA/Par\_file}
(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
solver.
\begin{itemize}
\item The adjoint simulation is launched together with the back reconstruction
of the original forward wavefield from the state variables saved from
the previous forward simulation, and the finite-frequency kernels
are computed by the interaction of the reconstructed forward wavefield
and the adjoint wavefield.
\item The back-reconstructed seismograms at the original station locations
are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
and can be collected to the local disk.
\item These back-constructed seismograms can be compared with the time-reversed
original seismograms to assess the accuracy of the backward reconstruction,
and they should match very well.
\item The arrays for density, P-wave speed and S-wave speed kernels are
also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
are the different types of kernels.
\end{itemize}
\item \textbf{Run the anisotropic kernel simulation}

Instead of the kernels for the isotropic wave speeds, you can also
compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
(using Voigt's notation) of the elastic tensor in the cartesian coordinate
system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
\texttt{.true.} in \texttt{DATA/Par\_file} before compiling the package.
The definition of the parameters $C_{IJ}$ in terms of the corresponding
components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
the processor number). The output kernels correspond to absolute perturbations
$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
For consistency, the output density kernels with this option turned
on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
anisotropic kernels can then be combined to obtain the kernels for
different parameterizations of anisotropy. This can be done, for example,
when combining the kernel files from slices into one mesh file (see
Section~\ref{sec:Finite-Frequency-Kernels}).

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{SAVE\_TRANSVERSE\_KL} \texttt{=} \texttt{.true.} in \texttt{DATA/Par\_file}
the package will save anisotropic kernels parameterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}

### 2022-11-11 doc/USER_MANUAL/09_noise_simulations.tex

With all of the above done, we can finally launch our simulations.
Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
is not globally shared. It is quite problematic to have a globally
shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
of I/O (we have to save the wavefields at the earth's surface at every
time step).

@@

After those simulations, you have all stuff you need, either in the
\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
in \texttt{DATA/Par\_file} (which are most probably on local nodes).
Collect whatever you want from the local nodes to your workstation,
and then visualize them. This process is the same as what you may
have done for regular earthquake simulations. Refer to other chapters
if you have problems.\\

>>>>

With all of the above done, we can finally launch our simulations.
Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{Par\_file}
is not globally shared. It is quite problematic to have a globally
shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
of I/O (we have to save the wavefields at the earth's surface at every
time step).

@@

After those simulations, you have all stuff you need, either in the
\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
in \texttt{Par\_file} (which are most probably on local nodes).
Collect whatever you want from the local nodes to your workstation,
and then visualize them. This process is the same as what you may
have done for regular earthquake simulations. Refer to other chapters
if you have problems.\\

### 2022-11-11 doc/USER_MANUAL/A_reference_frame.tex

The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
with respect to the surface of the mesh at the specified $x$/$y$
location to find the corresponding $z$ coordinate. It is possible
to use this depth value directly as $z$ coordinate by changing the
flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.

@@

The seismogram output directions are given in Cartesian $x$/$y$/$z$
directions and not rotated any further. Changing flags in \texttt{constants.h}
in the \texttt{src/shared/} subdirectory only rotates the seismogram
outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
set to \texttt{.false.}) and the normal to the surface at the receiver
location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
as vertical. In this case, the outputs are rotated to have the vertical
component normal to the surface of the mesh, $x$ and $y$ directions
are somewhat arbitrary orthogonal directions along the surface.

For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
Additionally, we add labels to the synthetics using the following
convention: For a receiver station located in an
\begin{description}
\item [{elastic domain:}] ~
\begin{itemize}
\item \texttt{semd} for the displacement vector
\item \texttt{semv} for the velocity vector
\item \texttt{sema} for the acceleration vector
\end{itemize}
\item [{acoustic domain:}] ~\\
 (please note that receiver stations in acoustic domains must be buried.
This is due to the free surface condition which enforces a zero pressure
at the surface.)
\begin{itemize}
\item \texttt{semd} for the displacement vector
\item \texttt{semv} for the velocity vector
\item \texttt{sema} for pressure which will be stored in the vertical component
\texttt{Z}. Note that pressure in the acoustic media is isotropic
and thus the pressure record would be the same in the other two component
directions. We therefore use the other two seismogram components to
store the acoustic potential in component \texttt{X} (or \texttt{N})
and the first time derivative of the acoustic potential in component
\texttt{Y} (or \texttt{E}).
\end{itemize}
\end{description}
The seismograms are by default written out in ASCII-format to the
\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
can change this behavior by changing the following flags in the \texttt{constants.h}
file located in the \texttt{src/shared/} subdirectory:
\begin{description}
\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
written out in binary format.
\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MAIN}}] Set to \texttt{.true.}
to have only the main process writing out seismograms. This can
be useful on a cluster, where only the main process node has access
to the output directory.
\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
have the seismograms output to \texttt{LOCAL\_PATH} directory specified
in the main parameter file \texttt{DATA/Par\_file}.
\end{description}

>>>>

The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
with respect to the surface of the mesh at the specified $x$/$y$
location to find the corresponding $z$ coordinate. It is possible
to use this depth value directly as $z$ coordinate by changing the
flag \texttt{USE\_SOURCES\_RECEIVERS\_Z} to \texttt{.true.} in
\texttt{DATA/Par\_file}.

@@

The seismogram output directions are given in Cartesian $x$/$y$/$z$
directions and not rotated any further. Changing flags in \texttt{setup/constants.h} only rotates the seismogram
outputs if receivers are forced to be located at the surface (\texttt{RECEIVERS\_CAN\_BE\_BURIED}
set to \texttt{.false.}) and the normal to the surface at the receiver
location should be used (\texttt{EXTERNAL\_MESH\_RECEIVERS\_NORMAL} set to \texttt{.true.})
as vertical. In this case, the outputs are rotated to have the vertical
component normal to the surface of the mesh, $x$ and $y$ directions
are somewhat arbitrary orthogonal directions along the surface.

For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
Additionally, we add labels to the synthetics using the following
convention: For a receiver station located in an
\begin{description}
\item [{elastic domain:}] ~
\begin{itemize}
\item \texttt{semd} for the displacement vector
\item \texttt{semv} for the velocity vector
\item \texttt{sema} for the acceleration vector
\end{itemize}
\item [{acoustic domain:}] ~\\
 (please note that receiver stations in acoustic domains must be buried.
This is due to the free surface condition which enforces a zero pressure
at the surface.)
\begin{itemize}
\item \texttt{semd} for the displacement vector
\item \texttt{semv} for the velocity vector
\item \texttt{sema} for pressure which will be stored in the vertical component
\texttt{Z}. Note that pressure in the acoustic media is isotropic
and thus the pressure record would be the same in the other two component
directions. We therefore use the other two seismogram components to
store the acoustic potential in component \texttt{X} (or \texttt{N})
and the first time derivative of the acoustic potential in component
\texttt{Y} (or \texttt{E}).
\end{itemize}
\end{description}
The seismograms are by default written out in ASCII-format to the
\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
can change this behavior by changing the following flags in the \texttt{DATA/Par\_file}:
\begin{description}
\item [{\texttt{USE\_BINARY\_FOR\_SEISMOGRAMS}}] set to \texttt{.true.} to have seismograms
written out in binary format.
\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MAIN}}] Set to \texttt{.true.}
to have only the main process writing out seismograms. This can
be useful on a cluster, where only the main process node has access
to the output directory.
\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
have the seismograms output to \texttt{LOCAL\_PATH} directory specified
in the main parameter file \texttt{DATA/Par\_file}.
\end{description}
